#Global trends for startup industries

### by Zhihui Xie

========================================================

## Global setting and library for this analysis

```{r global_options}
# global setting for this analysis
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8,
               warning=FALSE, message=FALSE)
```

```{r packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
```

# 1. Overview
This data set includes information about startups worldwide collected by CruchBase (https://info.crunchbase.com/about/crunchbase-data-exports/). The exploration here will focus on the global trends of startup industry occupied percentage over time from 1990 to 2013.

# 2. Clean data and Exploration

## 2.1 Load the Data

```{r Load_the_Data}
setwd("~/Downloads")
startups <- read.csv("crunchbase_monthly_export_companies_2014.csv")
```

## 2.2 Summary of the Data Set

```{r Summary}
dim(startups)
summary(startups)
```

## 2.3 Clean Data

```{r number of new founded industries}
# clean data: remove invalid data point in "market" and "founded_year" column
startups <- subset(startups, market != ""&!is.na(founded_year))
# calculate number of new founded startups in each maket in each year
startups.by.year.market <- startups %>%
  group_by(founded_year, market) %>%
  summarise(
            fre = n()
            )
```

## 2.4 First Exploration 

```{r first plot: include all founded years and industries}
p0 <- ggplot(aes(x = founded_year, y = fre), data = startups.by.year.market) +
  geom_line(aes(color = market)) + geom_point(aes(color = market))
p0
```

There is too much information there and it's hard to get the effective comparisons. Therefore, the data were truncated and focused on hot industries from 1990 to 2013.

## 2.5 Get a Subset of Data

```{r}
# limit data from 1990 to 2013
startups.by.year.market <- subset(startups.by.year.market, (1989 < founded_year) & (founded_year < 2014))
# find hot industries with total number of startups greater than 1400
markets <- as.data.frame(table(startups$market))
sub_markets <- subset(markets, Freq > 1400 & !Var1 == "") 
c_market <- c(as.character(sub_markets$Var1))
# list hot industrie for startups
c_market
hot_markets <- c(" Biotechnology ", " E-Commerce ", " Mobile ", " Software ")
# make table of hot industries and founded year
hot_startups.by.year.market <- subset(startups.by.year.market, market %in% hot_markets)
```

## 2.6 Second Exploration

```{r second plot: hot industries from 1990 to 2013}
p1 <- ggplot(aes(x = founded_year, y = fre), data = hot_startups.by.year.market) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  scale_x_continuous(breaks = seq(1990, 2013, 1)) +
  coord_cartesian(xlim = c(1990, 2014)) + 
  labs(x = "Year", y = "Number of Startups", color = "Industries") + 
  ggtitle("Global Trends for startup industries")
p1
```

## 2.7 Data Normalization

```{r reshape data}
# calculate total new founded startups each year
df0 <- startups %>%
  group_by(founded_year) %>%
  summarise(
            tol = n()
            )
# calculate new founded startups from 1990 to 2013 in each industry
df1 <- subset(df0, (1989 < founded_year) & (founded_year < 2014))

# merge hot startups and df1
df2 <- merge(hot_startups.by.year.market, df1, by = "founded_year", all = TRUE)
#head(df2)
# calculate percentage of startups in each industry each year
df3 <- df2 %>%
  group_by(founded_year, market) %>%
  summarise(
    percentage = fre/tol)
#head(df3)
# merge to obtain final subset of data
df.final <- merge(df2, df3, by = c("founded_year", "market"), all = TRUE)
#head(df.final)
```

## 2.8 Third Explorarion
```{r third plot: industry occupied percentage}
p2 <- ggplot(aes(x = founded_year, y = percentage), data = df.final) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  scale_x_continuous(breaks = seq(1990, 2013, 1)) +
  coord_cartesian(xlim = c(1990, 2014)) + 
  labs(x = "Year", y = "Percentage of Startups", color = "Industries") + 
  ggtitle("Trends for startup industries")
p2
```

It seems that all the selected hot induries for startups tend to decline. It's interesting to explore which industry for startups has a increased trend. To do that, number of startups greater than 300 in each industry were selected and visulized to find industry with increased trends. 

## 2.9 Explore industries with increased trends

```{r}
# add increased industries
sub_markets <- subset(markets, Freq > 300 & !Var1 == "") 
c_market <- c(as.character(sub_markets$Var1))
# plot each industry to find the one with increased trends.
# list hot industrie for startups with increased trends
other_markets <- c(" Real Estate ", " Travel ", " Fashion ", " Consulting ", " Education ", " Social Media ")
other_markets

startups.othermarket <- subset(startups.by.year.market, market %in% other_markets)
#df.other <- startups.othermarket %>%
#  group_by(founded_year) %>%
#  summarise(
#            market = " Other (Real Estate, Travel, Fashion, Consulting, Education, Social Media) ",
#            fre = sum(fre)
#            )
#hot_startups.by.year.market <- merge(hot_startups.by.year.market, df.other, by = c("founded_year", "market", "fre"), all = TRUE)

df2 <- merge(startups.othermarket, df1, by = "founded_year", all = TRUE)
#head(df2)
# calculate percentage of startups in each industry each year
df3 <- df2 %>%
  group_by(founded_year, market) %>%
  summarise(
    percentage = fre/tol)
#head(df3)
# merge to obtain final subset of data
df.other <- merge(df2, df3, by = c("founded_year", "market"), all = TRUE)
df.final <- merge(df.final, df.other, all = TRUE)
#head(df.final)
```


## 2.10 Fourth Explorarion
```{r fourth plot: final industry occupied percentage}
p3 <- ggplot(aes(x = founded_year, y = percentage), data = df.final) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  scale_x_continuous(breaks = seq(1990, 2013, 1)) +
  coord_cartesian(xlim = c(1990, 2014)) + 
  labs(x = "Year", y = "Percentage of Startups", color = "Industries") + 
  ggtitle("Trends for startup industries")
p3
```

## 2.11 Group data by time period

```{r time period data}
df.f1 <-subset(df.final, founded_year < 2000)
df.f1 <- df.f1 %>%
  group_by(market) %>%
  summarise (
    founded_year = "1990-1999",
    fre = sum(fre),
    tol = sum(tol),
    percentage = fre/tol
    )
df.f2 <- subset(df.final, founded_year > 1999)
df.f2 <- df.f2 %>%
  group_by(market) %>%
  summarise (
    founded_year = "2000-2013",
    fre = sum(fre),
    tol = sum(tol),
    percentage = fre/tol
    )
df <- merge(df.f2, df.f1, by = c("founded_year", "market", "fre", "tol", "percentage"), all = TRUE)
```

## 2.12 Fifth Exploration

```{r fifth plot: compare two time period}
p4 <- ggplot(aes(x = founded_year, y = percentage), data = df) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  labs(x = "Year", y = "Percentage of Startups", color = "Industries") + 
  ggtitle("Trends for startup industries")
p4
```

## 2.13 Export Data

```{r data exportation}
df.final1 <- df.final
colnames(df.final1) <- c("Founded year of startups", "Startup industry", "Number of startups in this industry", "Total number of founded startups", "Percentage of startups by industry")
colnames(df) <- c("Founded year of startups", "Startup industry", "Number of startups in this industry", "Total number of founded startups", "Percentage of startups by industry")
write.csv(df.final1, file="data.csv", row.names=FALSE)
write.csv(df, file="data1.csv", row.names=FALSE)
```
