#Global trends for startup markets

### by Zhihui Xie

========================================================

## Global setting and library for this analysis

```{r global_options}
# global setting for this analysis
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8,
               warning=FALSE, message=FALSE)
```

```{r packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
```

# 1. Overview
#### This data set includes information about startups worldwide collected by CruchBase (https://info.crunchbase.com/about/crunchbase-data-exports/). The exploration here will focus on the global trends of startup market share over time from 1990 to 2014.

# 2. Clean data and Exploration

## 2.1 Load the Data

```{r Load_the_Data}
setwd("~/Downloads")
startups <- read.csv("crunchbase_monthly_export_companies_2014.csv")
```

## 2.2 Summary of the Data Set

```{r Summary}
dim(startups)
summary(startups)
```

## 2.3 Clean Data

```{r number of new founded markets}
# clean data: remove invalid data point in "market" and "founded_year" column
startups <- subset(startups, market != ""&!is.na(founded_year))
# calculate number of new founded startups in each maket in each year
startups.by.year.market <- startups %>%
  group_by(founded_year, market) %>%
  summarise(
            fre = n()
            )
```

## 2.4 First Exploration 

```{r first plot: include all founded years and markets}
p0 <- ggplot(aes(x = founded_year, y = fre), data = startups.by.year.market) +
  geom_line(aes(color = market)) + geom_point(aes(color = market))
p0
```
##### There is too much information there and it's hard to get the effective comparisons. Therefore, the data were truncated and focused on hot markets from 1990 to 2014.

## 2.5 Get a Subset of Data

```{r}
# limit data from 1990 to 2014
startups.by.year.market <- subset(startups.by.year.market, (1989 < founded_year) & (founded_year < 2015))
# find hot markets
markets <- as.data.frame(table(startups$market))
sub_markets <- subset(markets, Freq > 1400 & !Var1 == "") 
c_market <- c(as.character(sub_markets$Var1))
# list hot market for startups
c_market
# add two interesting markets
markets <- c(" Biotechnology ", " E-Commerce ", " Mobile ", " Software ", " Social Media ", " Real Estate ")
# make table of hot markets and founded year
hot_startups.by.year.market <- subset(startups.by.year.market, market %in% markets)
```

## 2.6 Second Exploration

```{r second plot: hot markets from 1990 to 2014}
p1 <- ggplot(aes(x = founded_year, y = fre), data = hot_startups.by.year.market) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  scale_x_continuous(breaks = seq(1990, 2014, 1)) +
  coord_cartesian(xlim = c(1990, 2015)) + 
  labs(x = "Year", y = "Number of Startups") + 
  ggtitle("Global Trends for startup markets")
p1
```

## 2.7 Data Normalization

```{r reshape data}
# calculate total new founded startups each year
df0 <- startups %>%
  group_by(founded_year) %>%
  summarise(
            tol = n()
            )
# calculate new founded startups each year in each market
df1 <- startups %>%
  group_by(founded_year, market) %>%
  summarise(
            freq = n()
            )
# merge df0 and df1
df2 <- merge(df0, df1, by = "founded_year", all = TRUE)
#head(df2)
# calculate percentage of startups in each market each year
df3 <- df2 %>%
  group_by(founded_year, market) %>%
  summarise(
    freq = freq,
    tol = tol,
    percentage = freq/tol)
#head(df3)
# select final subset of data
df.final <- subset(df3, market %in% markets & (1989 < founded_year) & (founded_year < 2015))
#head(df.final)
```

## 2.8 Third Explorarion
```{r third plot: market share}
p2 <- ggplot(aes(x = founded_year, y = percentage), data = df.final) +
  geom_line(aes(color = market)) + geom_point(aes(color = market)) +
  scale_x_continuous(breaks = seq(1990, 2014, 1)) +
  coord_cartesian(xlim = c(1990, 2015)) + 
  labs(x = "Year", y = "Percentage of Startups") + 
  ggtitle("Global Trends for startup markets")
p2
```

## 2.9 Export Data

```{r data exportation}
write.csv(df.final, file="data.csv", row.names=FALSE)
```
